---
title: "Intro to R coding practice"
author: "Your name here"
date: "2/13/2025"
format: pdf
---

1.  Change your name in the YAML. Be sure to keep the quotation marks!

2.  Execute the following the code in the code chunk. In the space below the code chunk that reads **Answer**: describe in words (as specifically as possible) what the code is doing. What type of object is this?

```{r}
1:10
```

**Answer**:

3.  We will learn how to use the important function `sample()`, which *allows us to obtain a random sample in R*. Look at the Help file by typing `?sample` into your Console. After some experimenting, write code that randomly samples five numbers from the set of integers 6-10, *without* replacement. Note: it's okay if you get something different from those around you! We are *random*ly sampling after all!

```{r}

```

4.  In following code chunk, write code that randomly samples five numbers from the set of integers 6-10, *with* replacement.

```{r}

```

5.  In the following code chunk, write code to load in the `openintro` package. Then run the code in the code chunk. We will need this package to access the `cherry` data set in the next question.

```{r}

```

6.  In the Console, type in `?cherry` to open up the Help file for the `cherry` data set. What are the units for each of the variables?

**Answer**:

7.  Remember that rows in data frames represent observations. How many observations are there in the `cherry` data frame? Answer this *via code* by using the `nrow()` function and passing in the name of data frame of interest.

```{r}

```

Once you're finished, be sure to render one last time and submit the outputted PDF to the corresponding Gradescope assignment!
