---
title: "DataFest"
author: "Team name here!!"
format: pdf
---

```{r message = F}
# add extra packages as necessary
library(tidyverse)
library(readr)

# load data
```

## Introduction

Write your introduction here! Make sure to very clearly state the research question(s) you're trying to answer.

## Data cleaning/wrangling

Discuss at a high-level the wrangling you needed to do to get your data into a position to implement your methods below. (This might be one of the last parts of the report you right).

```{r}
# any code you might need to do macro-level wrangling
```

## Methods

Brief description of methods used to answer your research question(s), along with justification for why your methods is/are appropriate. Remember, we need at least one visualization and one simulation-based inference method!

There shouldn't any code in this section because you are just describing your methods.

## Results with interpretation

Implement your methods described above. This section is mostly code, with just some text for interpretation in context. 

```{r}

```

## Conclusion

Brief conclusion about what you would do differently/wish you had!

## Acknowledgements

Please briefly describe each team member's contribution to the project!
