---
title: "Randomization tests"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r message = F, include = F}
library(tidyverse)
library(kableExtra)
library(openintro)
```

## Test of independence

Participants in the study were 48 bank supervisors who identified as male and were attending a management institute at UNC in 1972. Each supervisor was asked to assume the role of personnel director of a bank.

-   Each person given a file to judge whether the person in the file should be promoted

-   The files were identical, except half of them indicated that the candidate was male, and the other half were indicated as female

-   Files were randomly assigned to bank managers

**Research question**: **Are individuals who identify their sex as female discriminated against in promotion decisions made by their managers who identify as male?**

#### Step 1

$H_{0}:$ \vspace{1cm}

$H_{A}:$ \vspace{0.5cm}

#### Step 2

```{r echo = F, message = F}
set.seed(1)
sex <- c(rep("male", 24), rep('female', 24))
decision <- c(rep("promote", 21), rep("not promote", 3), rep("promote", 14), rep("not promote", 10))
discrimination <- data.frame(sex, decision) |>
  sample_frac()
```


::::: {layout="[0.4, 0.6]"}
::: {#firstcol}
```{r message = F}
discrimination |> 
  slice(1:3)
```
:::

::: {#secondcol}
```{r echo = F, message = F}
ct <- discrimination |>
  group_by(sex, decision) |>
  summarise(n = n()) |>
  spread(decision, n) |>
  ungroup() |>
  mutate(total = `not promote`+ promote)

ct <- ct |>
  rbind(c("total", colSums(ct[-1])))

ct <- ct |> 
  kable() |>
  column_spec(1, border_right = T)
ct
```
:::
:::::

Obtain some relevant/useful summary statistics:

\vspace{2cm}

$\alpha=$

#### Step 3

Simulate under $H_{0}$:

-   Write down "promote" on \_\_\_\_\_\_ cards and "\_\_\_\_\_\_\_\_\_\_\_" on \_\_\_\_\_\_ cards.

-   Then repeat the following $B$ times:

    1.  Shuffle the cards

    2.  Deal out a stack of \_\_\_\_\_\_\_ to represent \_\_\_\_\_\_\_ candidates and \_\_\_\_\_\_\_\_ to represent \_\_\_\_\_\_\_ candidates.

    3.  

\vspace{1cm}

```{r echo = T}
#| code-line-numbers: true
set.seed(100)
n <- nrow(discrimination)
n_f <- sum(discrimination$sex == "female")
n_m <- sum(discrimination$sex == "male")
decisions <- discrimination$decision
B <- 1000
diff_props_null <- rep(NA, B)
for(b in 1:B){
  shuffled <- sample(decisions, n)
  rand_f <- shuffled[1:n_f]
  rand_m <- shuffled[-c(1:n_f)]
  
  p_f_sim <-  mean(rand_f == "promote")
  p_m_sim <-mean(rand_m == "promote")
  
  diff_props_null[b] <- p_f_sim - p_m_sim
}
```

\clearpage

```{r echo = F, fig.height=3}
null_dist <- data.frame(prop_diff = diff_props_null)
null_dist |>
  ggplot(aes(x = prop_diff)) +
  geom_histogram(binwidth = 0.05, col = "white") +
  labs(x = "Difference in promotion rates (female - male)",
       title = "Null distribution") +
  theme(text = element_text(size= 16)) +
  scale_x_continuous(n.breaks = 10)
```

-   p-value:

#### Step 4

-   Decision:

\vspace{1cm}

-   Conclusion:

\vspace{1cm}

-   Possible error:


\clearpage

## Test for difference in proportions

An experiment was conducted, consisting of two treatments on 90 patients who underwent CPR for a heart attack and subsequently went to the hospital. Each patient was randomly assigned to either:

- treatment group: received a blood thinner
- control group: did not receive a blood thinner

For each patient, the outcome recorded was whether they survived for at least 24 hours.

**Research question**: **Is the blood thinner treatment effective for patients who undergo CPR after a heart attack??**

#### Step 1

\vspace{2cm}


$H_{0}:$ \vspace{1cm}

$H_{A}:$ \vspace{0.5cm}

#### Step 2

```{r echo = F, message = F}
set.seed(1)
sex <- c(rep("male", 24), rep('female', 24))
decision <- c(rep("promote", 21), rep("not promote", 3), rep("promote", 14), rep("not promote", 10))
discrimination <- data.frame(sex, decision) |>
  sample_frac()
```


::::: {layout="[0.4, 0.6]"}
::: {#firstcol}
```{r message = F, eval = F}
cpr |>
  slice(1:3)
```
```{r echo = F, eval = T}
set.seed(1)
cpr |>
  sample_n(3) |>
  data.frame()
```
:::

::: {#secondcol}
```{r echo = F, message = F}
ct <- cpr |>
  mutate(group = as.character(group))|>
  group_by(group, outcome) |>
  summarise(n = n()) |>
  spread(outcome, n) |>
  ungroup() |>
  mutate(total = survived + died)

ct <- ct |>
  rbind(c("total", colSums(ct[-1])))

ct <- ct |> 
  kable() |>
  column_spec(1, border_right = T)
ct
```
:::
:::::

Obtain some relevant/useful summary statistics:

\vspace{2cm}

$\alpha=$

#### Step 3

Simulate under $H_{0}$:

-   Write down "\_\_\_\_\_\_\_\_\_\_\_"  on \_\_\_\_\_\_ cards and "\_\_\_\_\_\_\_\_\_\_\_" on \_\_\_\_\_\_ cards.

-   Then repeat the following $B$ times:

    1.  Shuffle the cards

    2.  Deal out a stack of \_\_\_\_\_\_\_ to represent \_\_\_\_\_\_\_ candidates and \_\_\_\_\_\_\_\_ to represent \_\_\_\_\_\_\_ candidates.

    3.  

\vspace{1cm}

```{r echo = T}
#| code-line-numbers: true
set.seed(310)
n_t <- sum(cpr$group == "treatment")
n_c <- sum(cpr$group == "control")
cards <- cpr$outcome
B <- 1000
diff_props_null <- rep(NA , B)
for(b in 1:B){
  shuffled <- sample(cards)
  treat_sim <- shuffled[1:n_t]
  control_sim <- shuffled[-c(1:n_t)]
  
  p_t_sim <- mean(treat_sim == "survived")
  p_c_sim <- mean(control_sim == "survived")
  
  diff_props_null[b] <- p_t_sim - p_c_sim
}
```

\clearpage

```{r echo = F, fig.height=3}
data.frame(diffs = diff_props_null) |>
  ggplot(aes(x = diffs))+
  geom_histogram(binwidth = 0.025, col = "white")+
  labs(x = "Difference in survival rates (treat - control)",
       title = "Null distribution") +
  theme(text = element_text(size= 16)) +
  scale_x_continuous(n.breaks = 10)
```

-   p-value:

#### Step 4

-   Decision:

\vspace{1cm}

-   Conclusion:

\vspace{1cm}

-   Possible error:
