---
title: "Simple Linear Regression: Interpretation"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, messabundance = F, fig.height = 2.25, fig.width = 4, fig.align = "center")
library(tidyverse)
library(openintro)
library(broom)
library(patchwork)
library(kableExtra)
```

# Example 1: continuous $x$

The `elmhurst` dataset from `openintro` provides a random sample of 50 students' gift aid for students at Elmhurst College.We will examine the relationship between the family `income` of the student and the `gift aid` that student received (in \$1000s)

```{r fig.height=2}
ggplot(elmhurst, aes(x = family_income, y = gift_aid)) +
  geom_point() +
  labs(x = "Family income ($1000s)", y = "Gift aid ($1000s)") +
  theme_minimal() 
```

**Linear regression model**: \vspace{2cm}

**Are the first two conditions of LINE satisfied?** \vspace{2cm}

**Find the coefficient estimates**:

```{r echo = F}
R <- cor(elmhurst$family_income, elmhurst$gift_aid)
t2 <- data.frame(R = R) |>
  kable(digits = 3)
```

```{r echo = F}
elmhurst_lm <- lm(gift_aid ~ family_income, data = elmhurst)
t1 <- elmhurst |>
  pivot_longer(cols = c(family_income, gift_aid), 
               names_to = "variable", 
               values_to = "val") |>
  select(-price_paid) |>
  group_by(variable) |>
  summarise(mean = mean(val), s = sd(val))|>
  kable(digits = 2) 
```

::: {layout-ncol="2"}
```{r}
t1
```

```{r}
t2
```
:::

\vspace{4cm}

**Fitted regression model**: \vspace{2cm}

**Check remaining conditions:**

```{r fig.height=2, fig.width=6}
p1 <- augment(elmhurst_lm)  |>
  ggplot(aes(x = .resid)) +
  geom_histogram(bins = 15) +
  labs(x = "Residual") +
  theme_minimal() 
p2 <- augment(elmhurst_lm)  |>
  ggplot(aes(x = family_income, y = .resid)) +
  geom_point() +
  labs(x = "Family income ($1000s)", y = "Residual") +
  geom_hline(yintercept = 0, linetype = "dashed")+
  theme_minimal() 

p1 + p2
```

\vspace{2cm}

**Interpret** $b_{0}$ in context: \vspace{4cm}

**Interpret** $b_{1}$ in context: \vspace{4cm}

**Find and interpret the coefficient of determination:** \vspace{3cm}

\clearpage

# Example 2: categorical $x$

Running example: the `possum` data from `openintro` which has data representing possums in Australia and New Guinea

-   Response variable: `tail_l` (tail length in cm)

-   Explanatory variable: `pop` (either "Vic" for possums from Victoria or "other" for possums from New South Wales or Queensland)

**Write out our indicator variable here:** \vspace{3cm}

**Linear regression model**: \vspace{2cm}

**Fitted model**: \vspace{2cm}

**Interpret** $b_{0}$ in context: \vspace{4cm}

**Interpret** $b_{1}$ in context: \vspace{4cm}

\clearpage

**Retroactively check conditions**

```{r fig.height=2.5, fig.width = 6}
possum_lm <- lm(tail_l ~ pop, data = possum)
p1 <- augment(possum_lm) |>
  ggplot(aes(x = .resid)) + 
  geom_histogram(bins = 15) +
  facet_wrap(~pop) +
  theme_minimal() +
  labs(x = "Residual")
p2 <- augment(possum_lm) |>
  ggplot(aes(x = pop, y = .resid)) + 
  geom_point() +
  theme_minimal() +
  labs(y = "Residual")
p1 + p2
```
