---
title: "Multiple Linear Regression (MLR)"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, message= F, messabundance = F, fig.height = 2.25, fig.width = 4, fig.align = "center")
library(tidyverse)
library(broom)
library(patchwork)
library(kableExtra)
library(readr)
parkinsons <- read_csv("https://raw.githubusercontent.com/midd-stat201-spring2025/midd-stat201-spring2025.github.io/refs/heads/main/data/parkinsons.csv")
pd <- parkinsons |>
  filter(avg.f.q == 1)
```

Want to understand what might help explain the voice `shimmer` of a patient with low vocal fundamental frequency.

```{r fig.width=6}
pd |>
  pivot_longer(cols = c(jitter, hnr), names_to = "variable", values_to = "val") |>
  ggplot(aes(x = val, y = shimmer, col = status )) +
  geom_point() +
  facet_wrap(~ variable, scales = "free_x") +
  theme_minimal() +
  labs(x = "Value")
```

## Model 1

**Multiple linear regression model:** \vspace{1.5cm}

```{r echo = T}
shimmer_lm <- lm(shimmer ~ hnr + jitter, data = pd) 
```

```{r}
tidy(shimmer_lm) |> 
 kable(digits = c(3,3,3,3,7))
```


**Fitted model**: \vspace{1.5cm}

**Interpretation of**:

-   **Intercept**: \vspace{3cm}

-   **Remaining coefficients**: \vspace{5cm}

**Model 1 fit**

```{r}
glance(shimmer_lm)|>
  select(r.squared, adj.r.squared) |>
  kable(digits = 4)
```

**Conditions**

```{r}
p2 <- augment(shimmer_lm) |>
  ggplot(aes(x= .resid)) + 
  geom_histogram(bins = 20) +
  theme_minimal() +
  labs(x = "Residual")

p3 <- augment(shimmer_lm) |>
  ggplot(aes(x=.fitted, y = .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Estimated shimmer", y = "Residual")
```

```{r fig.width=6}
p2 + p3
```


**Inference**

-   **Hypotheses**: \vspace{3cm}

-   **Which variables seem to be associated with `shimmer`? Why?** \vspace{3cm}

## Model 2

Let's see a model that now includes the `status` of the patient as a predictor:

```{r echo = T}
shimmer_lm2 <- lm(shimmer ~ hnr + jitter + status, data = pd)
```

```{r}
tidy(shimmer_lm2) |>
  kable(digits = c(3, 3, 3, 3, 7))
```

**Fitted model**: \vspace{1.5cm}

**Interpretation of**:

-   **Intercept**: \vspace{2cm}

-   **Remaining coefficients**: \vspace{5.5cm}

**Model 2 fit**

```{r}
glance(shimmer_lm) |>
  select(r.squared, adj.r.squared)|>
  kable(digits = 4)
```

**Which model seems "better", and why?**


