---
title: "Simple Linear Regression: Inference"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, message= F, messabundance = F, fig.height = 2.25, fig.width = 4, fig.align = "center")
library(tidyverse)
library(openintro)
library(broom)
library(patchwork)
library(kableExtra)
```

Data on 463 courses at UT Austin were obtained to answer the question: "What factors explain differences in instructor teaching evaluation scores?"

We will look at the variables:

-   `score`: course instructor's average teaching score, where average is calculated from all students in that course. Scores ranged from 1-5, with 1 being lowest.

-   `bty_avg`: course instructor's average "beauty" score, where average is calculated from six student evaluations of "beauty". Scores ranged from 1-10, with 1 being lowest.

**Linear regression model**: \vspace{1.5cm}

## Hypothesis test

**Hypotheses:** \vspace{3cm}

**Are conditions for SLR met?**

```{r fig.height=2.5, fig.width=10}
p1 <- ggplot(evals, mapping = aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Average beauty score", y = "Average rating")+
  geom_smooth(method = "lm", se = FALSE)+
  theme_minimal()
eval_mod <- lm(score ~ bty_avg, data = evals)
tidy_eval <- eval_mod |>
  tidy()
p2 <- eval_mod |>
  augment() |>
  ggplot(aes(x = .resid)) +
  geom_histogram(bins = 25) +
  labs(x = "Residual") +
  theme_minimal()

p3 <- eval_mod |>
  augment() |>
  ggplot(aes(x = bty_avg, y= .resid)) +
  geom_point()+
  labs(y = "Residual", x = "Average beauty score") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal()


p1 + p2 + p3
```

\vspace{4cm}

Assuming conditions met...

```{r}
tidy_eval |>
  kable(digits = c(3,3,3,3,5)) 
```

-   Fitted model: \vspace{1cm}

-   Interpretation of intercept and slope: \vspace{4.5cm}

**Calculate the test-statistic and p-value for the hypotheses:** \vspace{4cm}

::: {layout-ncol="2"}
**If instead** $H_{A}: \beta_{1} > 0$, what code would you write for the p-value?

**If instead** $H_{A}: \beta_{1} < 0$, what code would you write for the p-value?
:::

\vspace{1.5cm}

## Confidence Interval

Let's obtain a 95% CI for $\beta_{1}$. Assuming conditions for SLR met...

**Critical value code**: \vspace{1cm}

**Construct your interval**: \vspace{2cm}
