---
title: "Introduction to Simple Linear Regression"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, messabundance = F, fig.height = 2.25, fig.width = 4, fig.align = "center")
library(tidyverse)
library(openintro)
library(broom)
library(kableExtra)
```

Some simulated data that we might want to fit a linear regression model to:

```{r}
set.seed(2)
n <- 50
b0 <- 1
b1 <- 0.5
epsilon <- rnorm(n, 0, 1.5)
x <- sort(runif(n, -5,5))
y <- b0  + b1 * x + epsilon
fit <- coef(lm(y~x))

df <- data.frame(x, y)
p <- ggplot(df, aes(x=x, y = y)) +
  geom_point(size = 2) +
  theme_minimal() 
p
```

A line "fit" to the data above. What are the residuals of the highlighted observations?

```{r fig.width=5}
ids <- c(9, 22, 45)
x_fit <- x[ids]
y_hat <- fit[1] + fit[2] * x
y_fit <- y_hat[ids]
df_fit <- data.frame(x = x_fit, x_end = x_fit, y = y[ids], y_end = y_fit)
p1 <- p + geom_abline(slope = fit[2], intercept = fit[1], col = "yellow", linewidth = 2) +
  labs(title = "Option 1")
p_resid <- p1 + 
  geom_segment(data = df_fit,mapping= aes(x = x, xend = x_end, y = y, yend = y_end), linetype = "dashed", col = "#F8766D")+
    geom_point(data = df_fit, mapping = aes(x = x, y = y_end, col = "fitted"), shape = 4, size = 4)+
  geom_point(data = df_fit, mapping = aes(x = x, y = y, col = "observed"), size = 2) +
  theme(legend.title = element_blank(), plot.title = element_blank()) +
  labs(y = "y", x = "x")
p_resid
```

```{r}
df |>
  add_column(y_hat) |>
  mutate(residual = y - y_hat) |>
  ggplot(aes(x = x, y = residual)) +
  geom_point(size = 2) +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(data = df_fit, aes(x = x, y = y - y_end), size = 2, col = "#F8766D" ) +
  labs(title = "Residual plot", caption = "Red dots = specific points from previous plot")
```

\clearpage

# Running example

Using the `cherry` data from `openintro`: we will see if a linear regression model is appropriate to model the relationship between cherry tree volume (response) and diameter (explanatory).

**Linear regression model** (in context): \vspace{3cm}

::: {style="color: maroon"}
We check two conditions *before* fitting the the model.
:::

**Condition 1**: **L**inearity

```{r}
ggplot(cherry, aes(x= diam, y = volume)) +
  geom_point() +
  labs(title = "Cherry tree diameter vs. volume", x = "diameter (inches)", y = "volume (cubic feet)") +
  theme_minimal()
```

*Does the linearity condition appear to be met?* \vspace{2cm}

**Condition 2**: **I**ndependence

*Does the independence condition appear to be met?* \vspace{2cm}

**Fitted model** (in context): \vspace{3cm}

::: {style="color: maroon"}
*After* obtaining the fitted model, we have access to the residuals which we use to assess the remaining two conditions.
:::

**Condition 3**: **N**ormality

```{r}
cherry_lm <- lm(volume ~ diam, cherry)
augment(cherry_lm) |>
  ggplot(mapping = aes(x = .resid)) +
  geom_histogram(bins = 20) +
   theme_minimal() +
  labs(x = "residuals", title="Histogram of residuals for cherry model")
```

*Does the Normality condition appear to be met?*\vspace{2cm}

**Condition 4: E**qual variance

```{r}
augment(cherry_lm) |>
  ggplot(mapping = aes(x = diam, y = .resid)) +
  geom_point() +
   theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth =1, col ="darkgrey") +
  labs(y = "Residual", title = "Residual plot for cherry model", x = "diameter (inches)")
```

*Does the equal variance condition appear to be met?*\vspace{2cm}

**Given our checks of the conditions, should we believe the linear regression model is appropriate for these data?**
