---
title: "CI and HT for single mean via CLT"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, messabundance = F)
library(tidyverse)
library(openintro)
library(patchwork)
```

# Confidence interval for single mean

## Age at marriage (part 1)

```{r echo = F}
set.seed(2)
n <- 25
sd_age <- round(sd(age_at_mar$age), 2)
x <- sample(age_at_mar$age, n)
xbar <- mean(x)
sigma_age <- 5
```

In 2006-2010, the CDC conducted a thorough survey asking US women their age at first marriage. Suppose it is known that the standard deviation of the ages at first marriage is `r sigma_age` years. Suppose we randomly sample 25 US women and ask them their age at first marriage (plotted below). Their average age at marriage was `r xbar`.

```{r fig.height=2.5, fig.width=3}
ggplot(data.frame(x), aes(x=x)) +
  geom_histogram(binwidth = 2, col = "white") +
  labs(x = "Age at first marriage") +
  theme_minimal() 
```

**Obtain an 80% confidence interval for the mean age of US women at first marriage (not using simulation).**

Check conditions for CLT:

\vspace{3cm}

What does CLT tell us?

\vspace{2cm}

\clearpage

If conditions for CLT met, proceed to construct interval:

-   Point estimate: \vspace{1cm}

-   Standard error: \vspace{1cm}

-   Critical value (write code/draw picture):

\vspace{3cm}

Confidence interval: \vspace{1cm}

## Age at marriage (part 2)

```{r}
s <- sd(x)
```

Now suppose we don't know the population standard deviation, so all we have access to is the sample standard deviation of `r round(s,2)`.

**Obtain an 80% confidence interval for the mean age of US women at first marriage (not using simulation).**

Note: conditions are still met.

Proceed to construct interval:

-   Point estimate: \vspace{1cm}

-   Standard error: \vspace{1cm}

-   Critical value (write code/draw picture):

\vspace{3cm}

Confidence interval: \vspace{1cm}

\clearpage

# Hypothesis test for single mean

## Salinity example

```{r}
salinity_dat <- salinity
x <- salinity_dat$salinity_ppt
n <- length(x)
s <- sd(x)
df <- n-1
mu0 <- 38
```

We have `r n` salinity level measurements (ppt) collected from a random sample of water masses in the Bimini Lagoon, Bahamas (see below).

We want to test if the average salinity level in Bimini Lagoon is different from `r mu0` ppm at the $\alpha = 0.05$ level.

```{r fig.height=2.5, fig.width=3}
salinity_dat |>
  ggplot(aes(x = salinity_ppt)) +
  geom_histogram(bins = 10) +
  theme_minimal() +
  labs(x = "Salinity (ppt)") 
```

1.  Set hypotheses \vspace{3cm}
2.  Collect/summarise data, set $\alpha$ \vspace{3cm}
3.  Obtain null distribution, test statistic, and p-value

-   Check conditions for CLT: \vspace{3cm}

-   Can we proceed with CLT-based method?

-   Null distribution: \vspace{2cm}

-   Test-statistic: \vspace{3cm}

-   Distribution of test-statistic: \vspace{2cm}

-   p-value: draw picture and write code to obtain p-value \vspace{7cm}

4.  Make decision and conclusion in context.
