---
title: "Central Limit Theorem"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, messabundance = F)
library(tidyverse)
library(openintro)
library(patchwork)
```

# Sample means

## Height example

```{r}
mu <- round(mean(nba_heights$h_in), 2)
sigma <- round(sd(nba_heights$h_in), 2)
n <- 20
set.seed(11)
x <- sample(nba_heights$h_in, n)
p_nba <- ggplot(data.frame(x), aes(x = x)) +
  geom_histogram(bins = 20, color = "white") +
  labs(x = "Player height (in)") +
  theme_minimal() +
  theme(text = element_text(size = 13))
```

The average height of all NBA players in the 2008-9 season is `r mu` inches, with a population standard deviation of `r sigma` inches. We randomly sampled `r n` of these players and recorded their heights, as shown.

```{r fig.height=2, fig.width=3}
p_nba
```

***What is the sampling distribution of the sample mean heights?***

**Check conditions**

\vspace{4cm}

**Proceed if conditions met**

\vspace{2cm}

```{r fig.height=2.5, fig.width=4}
set.seed(11)
x <- sample(nba_heights$h_in, n)
sample_df <- data.frame(h_in = x, dist = "sample")
sampling_df <- data.frame(h_in = replicate(200, mean(sample(nba_heights$h_in, n))), dist = "sampling")
pop_df <- data.frame(h_in = nba_heights$h_in, dist = "population")

rbind(sample_df, sampling_df, pop_df) |>
  ggplot(aes(x = h_in, fill = dist, col = dist))+
  geom_histogram(bins = 25, alpha = 0.75,
                 mapping = aes(y = stat(density)))+
                 #position = "identity") +
  scale_fill_viridis_d(option = "F") +
  scale_color_viridis_d(option = "F") +
  theme_minimal() +
  labs(x = "Height (in)",
       title = "Different distributions",
       fill = NULL, col = NULL)
```

## Bank example

Customers are standing in line at a bank.

-   Let $X_{i}$ represent the service time for customer $i$.

-   Suppose that the average service time for all customers is 5 minutes, with a standard deviation of 6 minutes.

-   Assume that a bank currently has 36 customers in it, and all customers are independent of each other.

-   ***What is the probability that the average service time of all these customers is less than 4 minutes?***

**Check conditions**

\vspace{4cm}

**Proceed if conditions met**

\vspace{3cm}

\clearpage

# Sample proportions

## M&Ms

Mars, Inc. is the company that makes M&M's. In 2008, Mars changed their color distribution to have 13% red candies.

Let $\hat{p}$ represent the proportion of red M&M's in a random sample of $n$ M&M's.

***What is the sampling distribution of*** $\hat{p}$ ***if we take a random sample of size***:

```{r}
p <- 0.13
n <- 100
p1 <- ggplot() +
   geom_function(fun = dnorm, args = list(mean = p, sd = sqrt(p*(1-p)/n)))+
  theme_minimal() +
  scale_x_continuous(limits = c(-0.2, 1.1), n.breaks = 6) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
  labs(x = expression(hat(p)), title = expression(N(0.13, 0.034)),) +
  theme(text = element_text(size = 13))

n <- 10
p2 <- ggplot() +
   geom_function(fun = dnorm, args = list(mean = p, sd = sqrt(p*(1-p)/n)))+
  theme_minimal() +
  scale_x_continuous(limits = c(-0.5, 1.1), n.breaks = 6) +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
  labs(x = expression(hat(p)), title = expression(N(0.13, 0.106))) +
  theme(text = element_text(size = 13))
```

::: {layout-ncol="2"}
[$\mathbf{n = 100}$]{.underline}

[$\mathbf{n = 10}$]{.underline}

**Check conditions**

**Check conditions**

\vspace{5cm}

\vspace{5cm}

**Proceed if conditions met**

**Proceed if conditions met**

\vspace{5cm}

\vspace{5cm}

```{r fig.height=2, fig.width=2.5}
p1
```

```{r fig.height=2, fig.width=2.5}
p2
```
:::

\clearpage

# CLT-based CI for single proportion

$$\text{point est.} \pm \text{critical value} \times \text{SE}$$

## Poll example

A poll of 100 randomly sampled registered voters in a town was conducted, asking voters if they support legalized marijuana. It was found that 60% of respondents were in support.

***Find a 90% confidence interval for the true proportion of town residents in favor of legalized marijuana.***

**Check conditions**

\vspace{4cm}

**If conditions met, construct CI**

\vspace{8cm}

**Interpret the confidence interval in context!**
