---
title: "CI and HT for differences via CLT"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
      \usepackage{multicol}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}

```{r echo = F, message= F}
knitr::opts_chunk$set(echo = F, warning = F, messabundance = F)
library(tidyverse)
library(openintro)
library(patchwork)
library(kableExtra)
```

# Difference in proportions

A study in 1994 examined over 1400 randomly sampled dogs, some of which had been exposed to the herbicide 2,4-Dichlorophenoxyacetic acid. They wanted to do know if there is an increased risk in dogs in developing cancer due to exposure to 2,4-D.

```{r message = F}
cancer <- cancer_in_dogs |>
  mutate(group = if_else(order == "2,4-D", "herbicide", "control")) |>
  select(group, response)
ct <- cancer |>
  group_by(group, response) |>
  summarise(n = n()) |>
  spread(response, n) |>
  ungroup() |>
  mutate(total =  cancer + `no cancer`)

ct <- ct |>
  rbind(c("total", colSums(ct[-1])))

ct <- ct |> 
  kable() |>
  column_spec(1, border_right = T)
ct
```

### Confidence interval

Let population 1 be dogs exposed to 2,4-D, and population 2 be dogs not exposed to 2,4-D. We want a 95% CI for $p_{1} - p_{2}$, where $p_{i}$ is the rate of cancer in population $i$.

Obtain useful statistics: \vspace{3cm}

-   Check conditions for CLT: \vspace{5cm}

-   Are conditions for CLT met? \vspace{1cm}

-   Collect the components of confidence interval: \vspace{5cm}

-   Construct interval: \vspace{2cm}

-   Interpret: \vspace{3cm}

### Hypothesis test

Do the data provide strong evidence at the 0.05 level that the rate of cancer is higher for dogs exposed to 2,4-D than that of dogs not exposed to the herbicide?

-   Define hypotheses \vspace{3cm}

-   Obtain pooled proportion \vspace{3cm}

-   Check conditions for CLT \vspace{5cm}

-   Are conditions for CLT met?

-   What is the null distribution of $\hat{p}_{1} - \hat{p}_{2}$? \vspace{2cm}

-   Find the value of the test-statistic \vspace{2cm}

-   Draw picture and write code to obtain p-value \vspace{4cm}

\clearpage

## Difference in means

The Mauna Loa Observatory in Hawaii of monitors atmospheric solar, atmospheric, and meteorological parameters. We have data on annual atmospheric CO2 concentrations from 1980-2015. Specifically, we are interested in comparing CO2 levels between years 2000-2015 and years 1980-1999.

```{r fig.height = 2, message = F}
co2 <- read_csv("data/co2_annmean_mlo.csv") |>
  rename("CO2" = "mean")
co2_subset <- co2 |>
  filter(year %in% 1980:2015) |>
  mutate(group = if_else(year < 2000, "1980-1999", "2000-2015"))
ggplot(co2_subset,aes(x =CO2)) +
  geom_histogram(bins = 10) +
  facet_wrap(~group, scales = "free") +
  theme_minimal() +
  labs(x = "CO2 (ppm)")
co2_subset |>
  group_by(group) |>
  summarise(n = n(), xbar = mean(CO2), s = sd(CO2)) |>
  kable(digits = 2, col.names = c("years","n", "$\\bar{x}$", "s") )
```

### Confidence interval

Obtain a 90% confidence interval for the difference between the average atmospheric C02 levels (ppm) from years 2000-2015 and years 1980-1999.

-   Define parameters \vspace{2cm}

-   Check conditions for CLT \vspace{4cm}

-   Are conditions for CLT met? \vspace{0.5cm}

-   Collect the components of confidence interval: \vspace{4cm}

-   Construct interval: \vspace{2cm}

-   Interpret: \vspace{2cm}

### Hypothesis test

-   Define hypotheses \vspace{2cm}

-   We already checked conditions!

-   Find the value of the test-statistic and its distribution \vspace{5cm}

-   Draw picture and write code to obtain p-value \vspace{5cm}

-   What code would you write if $H_{A}$ was two-sided?
