---
title: "Simulation test for single mean"
format: 
  pdf:
    geometry:
      - top=2.5cm
      - left=2.5cm
      - right=2.5cm
      - bottom=2.5cm
include-in-header:
  - text: |
      \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

\vspace{-2cm}
## Research Question 1

Suppose I am interested in learning if Middlebury students who have an 8:15am class get at least 7.5 hours of sleep (on average). \vspace{1cm}

$H_{0}:$ \vspace{1cm}

$H_{A}:$ \vspace{1cm}

```{r echo = F, message= F}
library(tidyverse)
dat <- read_csv("../data/stat201_data.csv")
sleep <- dat$avg_sleep
n <- length(sleep)
xbar <- round(mean(sleep), 3)
set.seed(123)
```

Summary information: \vspace{0.75cm}

$\alpha:$ \vspace{0.5cm}

Start by obtaining a bootstrap distribution of sample means:

```{r echo = T, fig.height=3}
B <- 5000
boot_means <- rep(NA, B)
for(b in 1:B){
  x_boot <- sample(sleep, n, replace = T)
  boot_means[b] <- mean(x_boot)
}
```

```{r echo = F, fig.height=2}
p1 <- data.frame(means = boot_means) |>
  ggplot(aes(x = means)) +
  geom_histogram(col = "white", bins = 40) +
  labs(x = "Sample mean of average sleep per night (hours)", title = "Bootstrap distribution") +
  theme_minimal()+
  theme(text = element_text(size = 14))+
  scale_x_continuous(n.breaks = 8) 
p1
```

Why isn't this useful for obtaining p-value?


\clearpage

**How do we obtain the null distribution?**

```{r echo = T}
#| code-line-numbers: true
mu0 <- 7.5

# xbar stores observed sample mean
shift <- xbar - mu0

# boot_means is a vector holding B bootstrapped sample means
null_dist <- boot_means - shift
```

```{r echo = F, fig.height=2.5}
null_dist_df <- data.frame(means = null_dist) |>
  mutate(extreme = (means <= mu0 + shift) | (means >= mu0 - shift)  )
p2 <-  null_dist_df |>
  ggplot(aes(x = means)) +
  geom_histogram(col = "white", bins = 40) +
  labs(x = "Sample mean of average sleep per night (hours)", title = "Null distribution")  +
  theme_minimal()+
  theme(text = element_text(size = 14))+
  scale_x_continuous(n.breaks = 8) 
p2
```

p-value: \vspace{1cm}

Decision: \vspace{1cm}

Conclusion: \vspace{1cm}

\clearpage

## Research Question 2

Instead, I am interested in learning if Middlebury students who have an 8:15am class get 7.5 hours of sleep or not (on average). Then our hypotheses are:

\vspace{4cm}

How to find p-value now?

```{r echo = F, fig.height=2.5}
p2
```

\vspace{4cm}

p-value: \vspace{2cm}

Decision: \vspace{1cm}

Conclusion: \vspace{1cm}
