---
title: "Data wrangling with dplyr (part 1)"
format:
  html:
    toc: true
    toc-location: left
---

The textbook Modern Dive has some great descriptions and visuals to explain some of these functions. I highly recommend you read through the online textbook linked here: <https://moderndive.com/3-wrangling.html#filter>.

```{r message = FALSE}
library(tidyverse)
```

```{r eval = F}
books <- read_csv("data/amazon_books.csv", show_col_types = F)
```

```{r echo = F}
books <- read_csv("../data/amazon_books.csv",
                  show_col_types = F)

```

**By default, all `dplyr` functions expect the first argument to be a data frame.**

## Filter

In the following code, we use the `filter()` function to only retain the observations where the book's `list_price` is less than or equal to \$20, and then exactly equal to \$20. This function requires specifying a *logical condition*, and keeps observations in which the condition is met (i.e. TRUE).

```{r}
books |>
  filter(list_price <= 20)

books |>
  filter(amazon_price == 20)

books |>
  filter(is.na(height))
```

::: column-margin
What's going on here?

1.  Start with the data frame `books`
2.  Pipe (`|>`) the data frame to the `filter()` function and specify that we want to retain the cases where a particular condition is `TRUE`
3.  The result is a new data frame
:::

::: column-margin
We can use the following to create logical conditions: `==`, `>`, `<`, `>=`, `<=`, `is.na()`, `is.numeric()`, and many more!
:::

```{r}
books |>
  filter(is.na(height))
```

::: column-margin
To check for the lack of equality we use either `!=` or `!is.__()`. In general, the `!` reverses the boolean (so `!TRUE == FALSE`).
:::

::: {.callout-tip collapse="false" appearance="minimal"}
## Note

What is the difference between using `is.na()` within `filter()` and the following function `na.omit()`? Try running it!

```{r eval = F}
books |>
  na.omit()
```
:::

We can also filter for more than one condition at once. Within `filter()`, the comma `,` specifies that all conditions must be true. It can be read as "and". In the following code, we retain cases where a book's price on Amazon is that same as its list price, and has a list price under \$20.

```{r}
books |>
  filter(list_price == amazon_price, list_price < 20)
```

If we just need just one of multiple conditions to be true, we can use the `|` operator which stands for "or":

```{r}
books |>
  filter( (list_price == amazon_price) | list_price < 20)
```

::: column-margin
While not necessary, the extra `( )` for the first logical condition makes the code easier to read!
:::

We can also `filter()` based on categorical variables. If I only wanted hardcover books, I can write the following code:

```{r}
books |>
  filter(cover == "H")
```

If there are many levels of a categorical variable that I'd like to filter for, we can use `%in%` logical operator:

```{r}
books |>
  filter(publisher %in% c("Mariner Books", "Scribner"))
```

::: column-margin
I read this as: is this books's value of `publisher` *in* the vector `("Mariner Books", "Scribner")`?
:::

You could also use the "or" operator `|` but that can get a bit tedious!

```{r}
books |>
  filter( (publisher == "Mariner Books") | (publisher ==  "Scribner"))
```

## Mutate to add a new variable

It is typical for us to add new variables to a given data frame. We do this with the `mutate()` function. We must specify:

1.  The name of the new variable and
2.  How to calculate the value of that new variable *for each* observation. This will typically involve operations involving variables already present in the data frame. *Importantly, you should think of those function operating row-wise (i.e. each case will get its own new value).*

We link the two with ONE equals sign `=`.

In the following, we create a new variable called `body_wt_lbs` based off the rule `weight_oz / 16`. Thus, each case gets its own new value of `weight_lbs` based on its original value of `weight`. You can see the first three cases here:

```{r eval = F}
books |>
  mutate(weight_lbs = weight_oz / 16) 
```

```{r echo = F}
books |>
  mutate(weight_lbs = weight_oz / 16)  |>
  mutate(across(everything(), ~replace_na(as.character(.), "NA"))) |>
  slice(1:3) |>
  DT::datatable(options = list(pageLength = 3))
```

::: callout-tip
## Note

New variables from `mutate()` always get added to the last column of the data frame!
:::

### Storing our work from mutate

We're doing all this good data wrangling, but we're not "saving our work on the data frame". That is, even though we're using `dplyr` functions, all of the work is going into the ether unless we *deliberatively store our wrangled data frame*.

For example, what's the purpose of creating a new variable if we don't use it later on? Let's create a new data frame called `books2` that includes the new `weight_lbs` variable:

```{r}
books2 <- books |>
  mutate(weight_lbs = weight_oz / 16) 
```

::: callout-tip
## Note

Remember, when we store values in `R`, we never see the executed output. So when you run the above code, you don't see a snapshot of the data frame. If you wanted to see your new `books2` data frame, use `View()`!
:::

## Summarise

The `summarise()` function gives us an easy way to calculate summary statistics of variables in the data frame! We just need to know the name of the function such as `mean()` or `sd()` that will calculate the summary statistic for us.

Remember that the output is a data frame. Convention is to use the function like this: `summarise(informative_name = function for statistic)`.

Think about `summarise()` as operating column-wise: it will use ALL the cases for the specified variable.

Let's find the average weight (in pounds) of all our books:

```{r}
books2 |>
  summarise(mean_weight_lbs = mean(weight_lbs))
```

::: column-margin
Why did we use the data frame `books2` and not `books`?
:::

Oh no! We get an `NA` mean? That happens when some of the variables values themselves are `NA`. You can remove `NA`s using within `mean()` or `sd()` as shown below, OR you can `filter()` them out before getting to `summarise()`:

```{r}
books2 |>
  summarise(mean_weight_lbs = mean(weight_lbs, na.rm = T), 
            mean_amazon_price = mean(amazon_price))
```

::: column-margin
You can obtain multiple summary statistics at once by separating the desired summary statistics with commas.
:::

::: callout-tip
## Note

The `summarise()` function changes the data frame entirely. It collapses rows down to a summary statistic, and removes all columns that are irrelevant to the calculation.
:::
