---
title: "Numerical data coding practice"
author: "Your name"
date: "2/19/2025"
format: pdf
---

1.  Change your name in the YAML.

2.  In the following code chunk, load in the `openintro` package. Then run this code chunk. We need this package to once again work with the `cherry` data frame.

```{r}

```

3.  If we have a data frame (like `cherry`), we can access a specific variable's data in the data frame by typing the code like the following: `data_frame_name$variable_name`. Note that the spelling of the variable name must match how it appears in the data frame!

In the code chunk below, obtain the values of the observed diameters of these trees using code. Then answer: what type of object is output from the code?

```{r}


```

**Answer:**

4.  In the code chunk below, write code to find the mean and median of the height of trees in the `cherry` data frame. Based on what you find, answer the following: do you believe the distribution of the height of cherry trees is skewed or symmetric? Why?

```{r}

```

**Answer:**

5.  Let's confirm your answer above by creating a rather ugly (but easy to obtain) histogram. The `hist()` function requires one input: a vector of numerical values to visualize. Plot a histogram of the heights of these trees.

```{r}

```

6.  Optional: modify the plot above by specifying some of all of following extra inputs to the `hist()` function.

-   main: plot title (must be a string object)
-   xlab: x-axis label (must be a string object)
-   col: color to fill the bars in with (must be a valid color specified as a string object)

This looks like `hist(data, main = "Title", xlab = "x-axis title", col = "some color")`.

Once you're finished, be sure to render and submit the outputted PDF file to the corresponding Gradescope assignment!
