---
title: "Bootstrap confidence intervals"
author: "Your name"
date: "3/26/25"
format: pdf
---

We will work with the average hours of sleep our class reported. Run the following code chunk to load the data:

```{r message=FALSE}
library(tidyverse)
library(readr)
sleep <- read_csv("https://raw.githubusercontent.com/midd-stat201-spring2025/midd-stat201-spring2025.github.io/refs/heads/main/data/coffee_sleep.csv")
```

We will intialize a pseudo-random-number-generator using the `set.seed()` function. You can choose any whole number to input as the parameter to this function. For example, I used the seed of 201. If I have the exact same seed and code as you, then we will produce the same random results.

```{r}
set.seed(201)
```

1.  We will first take an original sample of size 10.

    Create a variable that represents and stores the target sample size. Then, take a random sample of size 10 of average sleep hours from our population. Store your sample into a variable called `x`.

```{r}
# create a variable for sample size

# obtain and store our sample

```

2.  Then, we will take 5000 bootstrap iterations. Store this value as a variable B for reproducibility.

    Then obtain a bootstrap distribution of the sample means using your original sample. *Remember to store the bootstrap statistics somewhere! Please use a meaningful variable name.* It may be useful to look at and modify the live code from previous class (on website). Your code shouldn't print out/show us any output.

```{r}
# store number of bootstrap iterations

```

3.  Use the `quantile()` function to obtain the bounds for a 80%, 90%, and 99% bootstrap confidence interval, respectively. Store these bounds as variables with different (but meaningful) variable names (you should have 6 variables total).

Your code shouldn't print out/show us any output.

```{r}
# 80%

# 90%

# 99%

```

4.  In the space provided below, report the three confidence intervals in the format of (lower, upper) *using in-line code*.

In-line code allows us to be reproducible. Remember, the format is: `r` where after the space within the back-ticks you type the appropriate R code.

80% CI:

90%: CI:

99% CI:

5.  How do the confidence interval widths change as the level of confidence increases?

**Answer:**

6.  Interpret one of your confidence intervals in context. *Use in-line code in your answer.*

**Answer:**
