---
title: "Coding practice: case_when"
author: "Your name"
format: pdf
---

0.  Change your name in the YAML and load in the packages necessary for wrangling. Then run the following chunk.

```{r message = F}
# add packages here
```

We will work with the `diamonds` data set again. Take a look at its Help file. When you're ready, create a new data frame called `diamonds2` that modifies `diamonds` to include an additional variable called `price_ct` which is the price of the diamond per carat

```{r}

```

2.  Create a new data frame that displays the proportion of diamonds whose price per carat of exceeds 8000 USD for each color of diamond. Your wrangled output should only display the color of the diamonds and the proportion that are greater than 8000 USD.

```{r}

```

3.  To your `diamonds2` data frame, use `case_when()` to add a variable called `carat_cat` that takes the value:

-   "< 1" if the diamond is less than 1 carat
-   "1-3" if the diamond is between 1 and 3 carats, inclusive
-   "> 3" if the diamond is more than 3 carats

Then using an appropriate plot type, visualize the distribution of price per carat for each of the categories you created. Make sure your plot has informative axis labels. What do you notice?

```{r}

```

**Answer:**

Once you're finished, render one last time and submit the rendered PDF to Gradescope!
