---
title: "Partner coding practice 2"
author: "Your names"
format: pdf
---

## Step 1: Load packages

In the code chunk below, write code to load in the `palmerpenguins` package and the package we use for wrangling. If you need to install the package, please do that first. We will once again work with the `penguins` data set.

```{r}

```

## Step 2: Focusing on particular groups

Create a new data frame called `adelie2009` that retains only the cases of Adelie penguins in the study year 2009. How many penguins are you left with?
 

```{r}

```

**Answer**:

## Step 3: Visualizing body mass

Using your new data frame, use an appropriate plot to visualize the distribution of body mass of these penguins across the different islands. Ensure that you have nice title and informative axis labels. 

```{r}

```

## Step 4: Adding more detail

To your code chunk above, bring in the sex of the penguin by either outlining the boxes in color by sex, or by filling them in with color by sex. (I know this is new to us; take a leap of faith!)

## Step 5: Numerical quantities

Still using your new data frame, provide a summary table/data frame that displays the mean body mass for all these penguins. Ensure that you have "nice" column names.

```{r}

```


## Step 6: Interpreting your EDA

Working together as a team, use the summary table and plot you both created to address the following questions:

1.  Which island has the biggest penguins? Does this depend on sex?

**Answer**:

2.  Do you think it is the best idea to report a single summary statistic for body mass? If not, what would you like to do to have a more informative summary table/data frame?

**Answer**:

