---
title: "Partner coding practice 1"
author: "Your names"
format: pdf
---

## Step 1: Install new package

We will visualize the `penguins` data from the `palmerpenguins` package. Please install the package now.

## Step 2: Load packages

In the code chunk below, write code to load in only the packages you will need to visualize the `penguins` data.

```{r}

```

## Step 3: Get to know the data

Answer the following: how many numerical and how many categorical variables are included in `penguins`? 

**Answer**: 

## Step 4: Visualize species

Create a visualization that visualizes the distribution of penguin species.

```{r}

```

## Step 5: Understanding bills

Create a new visualization that examines the relationship between bill length and bill depth of the penguins, colored by penguin species. 

```{r}

```

## Step 6: Understanding the effect of islands

Directly modifying the code in the above code chunk, add the location of the penguins ( island) to your visualization. There may be more than one method to address this, however, one method will more easily allow you to address the questions below!

## Step 7: Interpreting your plots

Working together as a team, use the plots you both created to address the following questions:

1.  Which species of penguin is represented the least in the `penguins` data set?

**Answer**:

2. Which species of penguin is only found on Biscoe Island?

**Answer**:

3. Suppose I make the statement: "Bill length and bill depth of these three species of penguins are strongly linearly associated.". What might you add to improve this statement?

**Answer**:
