---
title: "Partner Coding Practice 4"
author: "Your names"
format: pdf
---

We will work with the `satgpa` data from the `openintro` package. In the following code chunk, load in the `openintro` package and the package we use for wrangling. Then take a look at the Help file of the `satgpa` dataset.

```{r echo = F, message = F}
library(openintro)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, eval = F)
```

## Question 1

We'd like to know if performance on the SAT and high school GPA are correlated with first year GPA. Create a visualization that effectively shows the relationship between a student's 1) average SAT percentile (averaged across math and verbal), 2) their high school GPA, and 3) their first year GPA. Interpret what you find.

```{r}
satgpa |>
  mutate(sat_avg = (sat_v + sat_m)/2) |>
  ggplot(aes(x = sat_avg, y = fy_gpa, col = hs_gpa)) +
  geom_point() 
```

## Question 2

Using wrangling code, find the proportion of students who had better first year GPAs than high school GPAs.

```{r}
satgpa |>
  summarise(prop = mean(fy_gpa > hs_gpa))
```

## Question 3 
Create a new data frame called `satgpa2` that contains two new variables: one called `hs_gpa_cat` and another called `fy_gpa_cat`. Each of these variables should take the value "under" if the respective GPA is less than or equal to 3.0, and "over" if the respective GPA is greater than 3.0.


```{r}
satgpa2 <- satgpa |>
  mutate(hs_gpa_cat = if_else(hs_gpa <= 3, "under", "over"),
         fy_gpa_cat = if_else(fy_gpa <= 3, "under", "over"))
```

## Question 4

Using wrangling code, obtain the probability of having a first year GPA greater than 3.0, conditioning on the different levels of high school GPA category. Wrangle until your displayed output has only two rows and two variables. *In choosing how to reduce the number of rows and columns, think about what is the most important information you need to present.*

```{r}
satgpa2 |>
  group_by(hs_gpa_cat) |>
  count(fy_gpa_cat) |>
  mutate(cond_prop = n/sum(n)) |>
  filter(fy_gpa_cat == "over") |>
  select(hs_gpa_cat, cond_prop)
```
